### 2022-11-14 revision/review-response.Rmd
starting revision

---
output: 
  bookdown::word_document2:
    reference_docx: response_styles.docx
    number_sections: false
always_allow_html: true
urlcolor: blue
link-citations: true
csl: ../florida-scientist.csl
bibliography: ../refs.bib
---

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(message = F, echo = F, cache = F, warning = F)
```

```{r echo = F, cache = F, eval = F}
spelling::spell_check_files(here::here('revision/review-response.Rmd'))
```

We sincerely thank the reviewers for providing comments on our manuscript.  We provide a point-by-point response to these comments below. In general, most of the revisions address concerns...

All page and line numbers in the reviewer comments refer to the original submission.  Page numbers noted in our response refer to the revised submission.

## Reviewer 1

1) Does this paper contain new knowledge, or new interpretation of existing knowledge?

The authors developed a rapid-response dashboard for the purposes of monitoring the effects of the Piney Point wastewater release in Summer of 2021. The collaborative, multi-organizational approach to the response, data collection, and subsequent dissemination of important information concerning the release and its effects on the health of Tampa Bay is certainly a novel and important product.

2)  Does it duplicate in any substantial way material published elsewhere?

The paper does not duplicate material published. While data from the Piney Point response have been analyzed and published in other journals, this paper serves as a detailed roadmap for how the data were collected and wrangled into a useable, public-facing product.

3)  Are technical terms used and spelled correctly?

Yes, all technical terms are used and spelled correctly.

4) Are there major flaws in the study design, execution or analysis?

There are no major flaws in the methods used to produce the Piney Point response dashboard.

5)  Please list below any suggestions for improvement.

A brief review for grammar/spelling/punctuation may be warranted. For example, “e.g.” statements should never end in “etc” as the e.g. is intended for examples (e.g., page 6). I imagine that a copy-edit review prior to publication will catch most of these. However, the paper overall was well-written and structured in a logical, easy-to-follow manner.

Table 2 may need to be reformatted to adjust spacing (hanging ‘a’ in Pennsylvania).

* __Response__: Thank you for your comments and we are pleased you found the manuscript a valuable contribution.  We have proofed the revision for any grammar/spelling/punctuation issues, specifically addressing your comment about "e.g." statements.  The issue noted for Table 2 will be addressed during copy-editing as this is the unformatted version that will be changed on publication.

## Reviewer 2

<!-- # References -->